|

### 2020-01-01 appendix/appendix.Rmd
version 0.2

---
title: |
       | Supplementary Materials for "Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil"
author:
date: \today
abstract: "Does local oversight improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app prompts users to submit data about construction sites, sends such crowdsourced information to independent engineers, and contacts the mayors' offices about project delays. Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of bottom-up monitoring on public service performance and suggest that interventions targeted to other groups, or focusing on different issues, may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

@@

1. **Percentage of the project completed before the impact evaluation
	 started**: This is a column in the Ministry of Education dataset. The data
	 in this column were sampled in the beginning of the intervention.

2. **Percentage of the project reported as completed by the end of the
	 intervention period**: Another column in the Ministry of Education
	 dataset. The data were sampled six months after the intervention.

3. **Difference between the percentage reported as completed before and after
	 the intervention**: The difference between (1) and (2) above. Note
	 that according to @gerber2012field, this way to look into outcomes is
	 positive, as having a pre-treatment covariate that highly predicts the
	 outcome reduces the variance in the estimator considerably.

4. **Dummy indicator for finished constructions**: The Ministry of Education
	 dataset has a column with the construction status. We created a dummy to be
	 one when the construction was reported as finished and zero otherwise. The
	 data here were sampled six months after the intervention.

5. **Dummy indicator for cancelled constructions**: Similarly, we created a
	dummy variable from the Ministry of Education dataset indicating when the
	construction was reported as cancelled. The data here were sampled six months
	after the intervention.

6. **Number of schools where construction companies updated the conclusion
	 dates**: Lastly, the Ministry of Education dataset has a column with the
	 construction estimated completion date. We took the completion dated
	 reported before and six months after the intervention. If the dates
	 differed, we coded as one, and zero otherwise.

@@

1. Impressions
  - Mean: `r mean(fballt$impressoes)`
  - Total for all campaigns: `r sum(fballt$impressoes)`

2. App Download link clicks:
  - Mean: `r mean(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular'])`
  - Total for all campaigns: `r sum(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular'])`

@@

1. [Statista Total Facebook Users](https://www.statista.com/statistics/244936/number-of-facebook-users-in-brazil/)
2. [Statista Total Facebook Users by Age](https://www.statista.com/statistics/866282/facebook-user-share-brazil-age/)

>>>>

---
title: |
       | Supplementary Materials for "Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil"
author:
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
abstract: "Does local oversight improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app prompts users to submit data about construction sites, sends such crowdsourced information to independent engineers, and contacts the mayors' offices about project delays. Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of bottom-up monitoring on public service performance and suggest that interventions targeted to other groups, or focusing on different issues, may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

[^freire]: Postdoctoral Research Associate, The Political Theory Project, Brown
University, Providence, RI 02912, USA,
[`danilofreire@brown.edu`](mailto:danilofreire@brown.edu),
<http://danilofreire.github.io>, +1 (401) 584-2494. Corresponding author.

[^galdino]: Executive Director, Transparência Brasil, São Paulo, SP, Brazil,
[`mgaldino@transparencia.org.br`](mailto:mgaldino@transparencia.org.br),
<https://www.transparencia.org.br>.

[^mignozzetti]: School of International Relations, Fundação Getulio Vargas, São
Paulo, SP, Brazil and Wilf Family Department of Politics, NYU, NY, USA,
[`umberto.mig@nyu.edu`](mailto:umberto.mig@nyu.edu), <http://umbertomig.com>.

@@

1. Percentage of the project completed before the impact evaluation started:
	 This is a column in the Ministry of Education dataset. The data in this
	 column were sampled in the beginning of the intervention.

2. Percentage of the project reported as completed by the end of the
	 intervention period: Another column in the Ministry of Education dataset.
	 The data were sampled six months after the intervention.

3. Difference between the percentage reported as completed before and after
	 the intervention: The difference between (1) and (2) above. Note
	 that according to @gerber2012field, this way to look into outcomes is
	 positive, as having a pre-treatment covariate that highly predicts the
	 outcome reduces the variance in the estimator considerably.

4. Dummy indicator for finished constructions: The Ministry of Education
	 dataset has a column with the construction status. We created a dummy to be
	 one when the construction was reported as finished and zero otherwise. The
	 data here were sampled six months after the intervention.

5. Dummy indicator for cancelled constructions: Similarly, we created a
	dummy variable from the Ministry of Education dataset indicating when the
	construction was reported as cancelled. The data here were sampled six months
	after the intervention.

6. Number of schools where construction companies updated the conclusion
	 dates: Lastly, the Ministry of Education dataset has a column with the
	 construction estimated completion date. We took the completion dated
	 reported before and six months after the intervention. If the dates
	 differed, we coded as one, and zero otherwise.

@@

1. Impressions
- Mean: `r format(round(mean(fballt$impressoes),2), big.mark = ",")`
	- Total for all campaigns: `r format(sum(fballt$impressoes), big.mark = ",")`

2. App Download link clicks:
	- Mean: `r format(mean(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular']), big.mark = ",")`
  - Total for all campaigns: `r format(sum(fballt$impressoes[fballt$tipo_de_resultado=='Instalações do aplicativo para celular']), big.mark = ",")`

@@

1. [Statista Total Facebook
	 Users](https://www.statista.com/statistics/244936/number-of-facebook-users-in-brazil/)

2. [Statista Total Facebook Users by
	 Age](https://www.statista.com/statistics/866282/facebook-user-share-brazil-age/)

### 2020-01-01 article/article.Rmd
version 0.2

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author:
date: \today
abstract: "Does ~~grassroots participation~~ **local oversight** improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. ~~The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact the mayor's office about project delays.~~ **The app prompts users to submit data about construction sites, sends such crowdsourced information to independent engineers, and contacts the mayors' offices about project delays.** Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of ~~local~~ **bottom-up** monitoring on public service performance and suggest that interventions targeted ~~at other groups~~ **to other groups, or focusing on different issues,** may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin:
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    template: article-template.latex
---

@@

@@

A robust accountability system is crucial for efficient public services
provision [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent;
@o1998horizontal]. In its standard definition, accountability is understood as
the process of holding authorities responsible for their actions
[@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past
studies show that accountability has a positive impact on governance; ~~as~~ it
ensures that politicians act on behalf of voters [@freire2010ngp;
@moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking
and corruption [@deininger2005does; @wenar2006accountability], and improves the
quality of public services [@adsera2003you; @bjorkman2009power]. Recent
research also suggests that accountability leads to higher economic growth
because it limits state discretion in the economy and increases long-term
investments in human capital [@benhabib2010economic; @suebvises2018social;
@ponzetto2018social].

However, accountability ~~systems take many forms~~ **mechanisms extend beyond
elections.** One promising model is that of _bottom-up monitoring_, in which
citizens receive information about the shortcomings of a given project so they
can evaluate and pressure underperforming public officials [@kosack2014does;
@molina2016community; @raffler2018weakness]. Proponents argue that bottom-up
accountability is effective because: 1) constituents have first-hand
information about the outcomes of local policies; 2) citizens have incentives
to attack corruption that directly affects themselves; 3) policy-makers are
sensitive to social punishment from their own communities [@serra2011combining,
570]. In this regard, bottom-up accountability offers a potential solution to
the principal-agent dilemma in public service by aligning the interests of
state officials with those of the constituency they serve [@barro1973control;
@raffler2018weakness].

@@

We use the TDP app to conduct two experimental interventions and test ~~its
impact~~ **the impact of citizen oversight** on five outcomes related to school
completion rates and complaints to public authorities. Overall, ~~our results
show that~~ providing information to citizens has no consistent impact on
policy outcomes. In the first experiment, we find that the TDP app increased
the likelihood of construction cancellation by 2 percent, but the result does
not replicate. The remaining five models have null results. In our second
intervention, none of the estimations reach conventional levels of statistical
significance. Importantly, all coefficients are small, what suggests that even
if the TDP app had a significant effect on the outcomes, its substantive impact
would be ~~virtually~~ negligible.

The findings raise questions about the ability of citizens to hold
representatives accountable using bottom-up ~~strategies~~ **monitoring**. 
\sout{Although studies like} **On the one hand,** @bjorkman2009power,
@lagunes2018guardians, and @reinikka2005fighting report better policy outcomes
after providing information to local communities. ~~our findings are closer
to~~ **On the other hand,** @banerjee2010pitfalls, @keefer2014mass,
@lieberman2014does, @bjorkman2017experimental, @olken2007monitoring, and
@raffler2018weakness, find little evidence that information-based interventions
lead to greater government responsiveness. **Our results are in line with the
latter group and suggest that local oversight is ineffective in altering
government behaviour in Brazil.**

@@

Between August 2017 and July 2019, we implemented two interventions to measure
the effect of the TDP app on five school construction outcomes plus a placebo
test. The outcomes are: 1) a placebo outcome indicating the percentage of the
project completed before the impact evaluation started; 2) the percentage of
the project completed by the end of the interventions; 3) the difference
between the percentage reported as completed before and after the
interventions; 4) the number of finished constructions; 5) the number of
cancelled constructions; 6) the number of schools where construction companies
updated the conclusion dates. ~~In terms of expectations,~~ Table
\ref{tab:outcomes} depicts the expected effects for each of the studied
outcomes[^more-details].

[^more-details]: **Please refer to the Online Appendix for further details on
the treatment implementation and the coding of the outcome variables.**

@@

We also conducted two manipulation checks and analysed the number of TDP app
downloads by municipality and over time. Figure \ref{fig:manipulation1}
displays the results and indicates that the treatment has good territorial
variability. There are `r length(unique(ga$municipality[ga$phase1==1]))`
downloads in the 
`r format(length(unique(impact_evaluation_phase1$municipality[impact_evaluation_phase1$treat=='Treatment'])),
big.mark = ",")` municipalities in the treatment condition. Downloads peak
during the Facebook TDP campaign, launched in October 2017, then diminish in
the following months. **Overall, 
`r format(sum(ga$novos_usuarios[ga$phase1==1]), big.mark = ",")` users
downloaded the app during intervention 1.**

```{r man1, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=4, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation1}Manipulation checks for intervention 1. The first plot shows the geographical distribution of the treatment condition and the second graph displays the number of TDP app downloads from August 2017 to July 2018.'}

@@

The second intervention is similar to intervention 1 in all but three
characteristics. First, the TDP app was **then** available for both Android and
iOS devices. Second, we randomised the intervention at the school level, with
`r as.numeric(table(impact_evaluation_phase2$treat)[1])` control and 
`r format(as.numeric(table(impact_evaluation_phase2$treat)[2]), big.mark = ",",
scientific = FALSE)` treatment units. We used blocked randomisation stratified
by Brazilian states, school construction status (under construction, stopped,
unfinished), and whether the municipality spent more on school construction
that the distribution median. Finally, the intervention period lasted from
August 2018 to July 2019.

Balance tests and manipulation checks were also successful for intervention 2.
In total, `r length(unique(ga$municipality[ga$phase2==1]))` municipalities
downloaded the app. There is about 1,000 user downloads in August 2018, right
after intervention 2 starts, and a second spike around December. 
**The app gained 
`r format(sum(ga$novos_usuarios[ga$phase1==0]), big.mark = ",")` new users
during intervention 2. The number of downloads is smaller in this second
intervention as there was no associated social media campaign in that period.**

@@

**Data from Google Analytics suggest that users did engage with the TDP app. On
average, each user launched `r format(round(mean(as.numeric(gsub('\\.', '',
ga2$visualizacoes_de_tela))), 1), big.mark = ",")` app sessions, which
indicates their interest in the application. In total, the app had 
`r format(sum(as.numeric(gsub('\\.', '', ga2$visualizacoes_de_tela))),
big.mark = ",", scientific = F)` screen visualisations, with an average of 
`r round(mean(as.numeric(gsub(',', '.', ga2$telas_sessao))), 2)` screen
visualisations per session.**

@@

We find that the app only has a small effect on cancellation rates. The TDP
application increases the likelihood of cancelling the construction in 
`r round(100*resInt1[[2]][[20]]$coefficients[1], 2)` percent. While this result
goes in the opposite direction of our theoretical expectations, the finding is
inconsistent and does not replicate in the second experiment. All other
coefficients are not statistically significant at conventional levels. On the
one hand, the results indicate that our placebo outcome, the percentage of the
invested executed before the intervention, indeed behaves as predicted. On the
other hand, we expected the five remaining outcomes to improve after the
introduction of the app. The literature on bottom-up accountability argues that
delivering more information about the shortcomings of public services provision
will ~~mobilise citizens who, in turn, would pressure state agents to produce
better social policies~~ **put citizens in a position where they can monitor
state agents and improve provider behaviour** [@raffler2018weakness]. Our
results do not lend support to that hypothesis.

@@

We also note that effect sizes are small in all estimations and that the
coefficients flip signs in all but one of our six dependent variables. This
provides further evidence for the null results: not only the app would have
a low impact even if the treatment were significant, but the results could go
against what the bottom-up accountability theory predicts. At least in the
school construction outcomes we investigate here, we find little evidence that
grassroots monitoring works in the context of school constructions in Brazil.

@@

In this paper, we discuss **whether delivering information to citizens via** a
mobile phone application fosters ~~grassroots mobilisation~~ **community
oversight** and political accountability in Brazil. Our two interventions show
that the results are at best mixed. Although we find ~~an~~ **some treatment**
effect on school cancelling rates in the first intervention, the app has no
consistent impact on our outcomes of interest. These findings add to the
studies that cast doubts on the relationship between bottom-up accountability
and local policy performance [e.g., @banerjee2010pitfalls; @lieberman2014does;
@raffler2018weakness].

What factors, then, are driving these results? It seems unlikely that the null
results derive from flaws in the research design. First, our study is well
powered. Although the treatment is indirect---the person has to download the
app, find a school construction, and then report it---,we included
a substantial number of schools in the treatment groups. Second, balance and
manipulation tests indicate that the treatment allocation was successful, so we
can rule out problems in the randomisation procedures. **Third, data from
Google Analytics confirm that citizens indeed used the app and provided
information to our dataset. This indicates that the treatment manipulation was
effective. In this sense, it is unlikely that our results derive from low user
response.** Fourth, after doing a series of robustness tests, we still find no
firm evidence of treatment effect. Finally, note that the sign of the
coefficients are frequently contrary to our theoretical expectations. This
rules out a possible concern about statistical power with our small control
group approach.

**We discuss some possible reasons why community monitoring did not work in our
case**. ~~A~~ **One** plausible explanation is that individuals were unable to
differentiate the effect of political corruption from those of spending cuts.
Due to the severe economic crisis in 2014--2016, the Brazilian federal
government introduced discretionary spending limits that affected public
investment [@rossi2016impactos]. Politicians may argue that
delays in school constructions are not derived from their misuse of government
funds, but from the austerity measures. If this is the case, citizens will not
blame local politicians for the underprovision of public goods. Consequently,
representatives can dismiss individual requests as the issue is unlikely to
escalate.

**The electoral cycle might also have decreased the potential effect of the
treatment. As the experiment was fielded right after Brazil's municipal
elections, incumbents might have disregarded the requests because they did not
see the demands as politically costly in the short run. Having just taken
office, mayors might have focused their attention to the formation of
government coalitions or to budget concerns. Future research may evaluate how
electoral dynamics interact with citizen oversight, potentially by replicating
informational experiments in different stages of the political cycle.**

\sout{Moreover, it is possible that a mobile app like TDP is simply not
sufficient to promote collective action. Raffler et al (2018, 29) point
out several reasons why information interventions does not lead to higher
pressure on public agents, such as citizens' apathy towards their ability to
make public agents accountable, lack of official channels to support popular
demands, or that constituents have more pressing needs than monitoring
representatives. Furthermore, individuals may also believe they are better off
by free-riding on other users of the TDP app, thus reaping the benefits of
having faster school constructions without incurring personal costs (Olson
1965).}

In sum, our experiments suggest that popular participation and bottom-up
monitoring may not be effective to improve public service delivery in the case
of school constructions in Brazil. Nevertheless, the null findings are
informative to researchers and policy-makers. The most important recommendation
derived from this study is that interventions targeting elite groups, such as
lobbyists or civil servants, might render better ~~policy~~ **school
construction** outcomes than those focused at the ~~grassroots levels~~
**community level**. Another core lesson is that although digital interventions
are promising means to deliver information, perhaps they do not have the same
impact as personal, face-to-face communication. Since many developing countries
share Brazil's issues with education provision, the shortcomings we describe
here serve as warnings for future interventions. Finally, whether this study
generalises beyond school constructions to other bottom-up programs, and to
other contexts, remains to be studied.

\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}

>>>>

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author:
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
abstract: "Does local oversight improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app prompts users to submit data about construction sites, sends such crowdsourced information to independent engineers, and contacts the mayors' offices about project delays. Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of bottom-up monitoring on public service performance and suggest that interventions targeted to other groups, or focusing on different issues, may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin:
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    template: article-template.latex
---

@@

[^freire]: Postdoctoral Research Associate, The Political Theory Project, Brown
University, Providence, RI 02912, USA,
[`danilofreire@brown.edu`](mailto:danilofreire@brown.edu),
<http://danilofreire.github.io>, +1 (401) 584-2494. Corresponding author.

[^galdino]: Executive Director, Transparência Brasil, São Paulo, SP, Brazil,
[`mgaldino@transparencia.org.br`](mailto:mgaldino@transparencia.org.br),
<https://www.transparencia.org.br>.

[^mignozzetti]: School of International Relations, Fundação Getulio Vargas, São
Paulo, SP, Brazil and Wilf Family Department of Politics, NYU, NY, USA,
[`umberto.mig@nyu.edu`](mailto:umberto.mig@nyu.edu), <http://umbertomig.com>.

@@

A robust accountability system is crucial for efficient public services
provision [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent;
@o1998horizontal]. In its standard definition, accountability is understood as
the process of holding authorities responsible for their actions
[@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past
studies show that accountability has a positive impact on governance; it
ensures that politicians act on behalf of voters [@freire2010ngp;
@moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking
and corruption [@deininger2005does; @wenar2006accountability], and improves the
quality of public services [@adsera2003you; @bjorkman2009power]. Recent
research also suggests that accountability leads to higher economic growth
because it limits state discretion in the economy and increases long-term
investments in human capital [@benhabib2010economic; @suebvises2018social;
@ponzetto2018social].

However, accountability mechanisms extend beyond elections. One promising model
is that of _bottom-up monitoring_, in which citizens receive information about
the shortcomings of a given project so they can evaluate and pressure
underperforming public officials [@kosack2014does; @molina2016community;
@raffler2018weakness]. Proponents argue that bottom-up accountability is
effective because: 1) constituents have first-hand information about the
outcomes of local policies; 2) citizens have incentives to attack corruption
that directly affects themselves; 3) policy-makers are sensitive to social
punishment from their own communities [@serra2011combining, 570]. In this
regard, bottom-up accountability offers a potential solution to the
principal-agent dilemma in public service by aligning the interests of state
officials with those of the constituency they serve [@barro1973control;
@raffler2018weakness].

@@

We use the TDP app to conduct two experimental interventions and test the
impact of citizen oversight on five outcomes related to school completion rates
and complaints to public authorities. Overall, providing information to
citizens has no consistent impact on policy outcomes. In the first experiment,
we find that the TDP app increased the likelihood of construction cancellation
by 2 percent, but the result does not replicate. The remaining five models have
null results. In our second intervention, none of the estimations reach
conventional levels of statistical significance. Importantly, all coefficients
are small, what suggests that even if the TDP app had a significant effect on
the outcomes, its substantive impact would be negligible.

The findings raise questions about the ability of citizens to hold
representatives accountable using bottom-up monitoring. On the one hand,
@bjorkman2009power, @lagunes2018guardians, and @reinikka2005fighting report
better policy outcomes after providing information to local communities. On the
other hand, @banerjee2010pitfalls, @keefer2014mass, @lieberman2014does,
@bjorkman2017experimental, @olken2007monitoring, and @raffler2018weakness, find
little evidence that information-based interventions lead to greater government
responsiveness. Our results are in line with the latter group and suggest that
local oversight is ineffective in altering government behaviour in Brazil.

@@

Between August 2017 and July 2019, we implemented two interventions to measure
the effect of the TDP app on five school construction outcomes plus a placebo
test. The outcomes are: 1) a placebo outcome indicating the percentage of the
project completed before the impact evaluation started; 2) the percentage of
the project completed by the end of the interventions; 3) the difference
between the percentage reported as completed before and after the
interventions; 4) the number of finished constructions; 5) the number of
cancelled constructions; 6) the number of schools where construction companies
updated the conclusion dates. Table \ref{tab:outcomes} depicts the expected
effects for each of the studied outcomes[^more-details].

[^more-details]: Please refer to the Online Appendix for further details on the
treatment implementation and the coding of the outcome variables.

@@

We also conducted two manipulation checks and analysed the number of TDP app
downloads by municipality and over time. Figure \ref{fig:manipulation1}
displays the results and indicates that the treatment has good territorial
variability. There are `r length(unique(ga$municipality[ga$phase1==1]))`
downloads in the 
`r format(length(unique(impact_evaluation_phase1$municipality[impact_evaluation_phase1$treat=='Treatment'])),
big.mark = ",")` municipalities in the treatment condition. Downloads peak
during the Facebook TDP campaign, launched in October 2017, then diminish in
the following months. Overall, 
`r format(sum(ga$novos_usuarios[ga$phase1==1]), big.mark = ",")` users
downloaded the app during intervention 1.

```{r man1, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=3.8, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation1}Manipulation checks for intervention 1. The first plot shows the geographical distribution of the treatment condition and the second graph displays the number of TDP app downloads from August 2017 to July 2018.'}

@@

The second intervention is similar to intervention 1 in all but three
characteristics. First, the TDP app was then available for both Android and iOS
devices. Second, we randomised the intervention at the school level, with 
`r as.numeric(table(impact_evaluation_phase2$treat)[1])` control and 
`r format(as.numeric(table(impact_evaluation_phase2$treat)[2]), big.mark = ",",
scientific = FALSE)` treatment units. We used blocked randomisation stratified
by Brazilian states, school construction status (under construction, stopped,
unfinished), and whether the municipality spent more on school construction
that the distribution median. Finally, the intervention period lasted from
August 2018 to July 2019.

Balance tests and manipulation checks were also successful for intervention 2.
In total, `r length(unique(ga$municipality[ga$phase2==1]))` municipalities
downloaded the app. There is about 1,000 user downloads in August 2018, right
after intervention 2 starts, and a second spike around December. 
The app gained 
`r format(sum(ga$novos_usuarios[ga$phase1==0]), big.mark = ",")` new users
during intervention 2. The number of downloads is smaller in this second
intervention as there was no associated social media campaign in that period.

@@

Data from Google Analytics suggest that users did engage with the TDP app. On
average, each user launched `r format(round(mean(as.numeric(gsub('\\.', '',
ga2$visualizacoes_de_tela))), 1), big.mark = ",")` app sessions, which
indicates their interest in the application. In total, the app had 
`r format(sum(as.numeric(gsub('\\.', '', ga2$visualizacoes_de_tela))),
big.mark = ",", scientific = F)` screen visualisations, with an average of 
`r round(mean(as.numeric(gsub(',', '.', ga2$telas_sessao))), 2)` screen
visualisations per session.

@@

We find that the app only has a small effect on cancellation rates. The TDP
application increases the likelihood of cancelling the construction in 
`r round(100*resInt1[[2]][[20]]$coefficients[1], 2)` percent. While this result
goes in the opposite direction of our theoretical expectations, the finding is
inconsistent and does not replicate in the second experiment. All other
coefficients are not statistically significant at conventional levels. On the
one hand, the results indicate that our placebo outcome, the percentage of the
invested executed before the intervention, indeed behaves as predicted. On the
other hand, we expected the five remaining outcomes to improve after the
introduction of the app. The literature on bottom-up accountability argues that
delivering more information about the shortcomings of public services provision
will put citizens in a position where they can monitor state agents and improve
provider behaviour [@raffler2018weakness]. Our results do not lend support to
that hypothesis.

@@

We also note that effect sizes are small in all estimations and that the
coefficients flip signs in all but one of our six dependent variables. This
provides further evidence for the null results: not only the app would have
a low impact even if the treatment were significant, but the results could go
against what the bottom-up accountability theory predicts. At least in the
school construction outcomes we investigate here, we find little evidence that
local monitoring works in the context of school constructions in Brazil.

@@

In this paper, we discuss a mobile phone application fosters community
oversight and political accountability in Brazil. Our two interventions show
that the results are at best mixed. Although we find some treatment effect on
school cancelling rates in the first intervention, the app has no consistent
impact on our outcomes of interest. These findings add to the studies that cast
doubts on the relationship between bottom-up accountability and local policy
performance [e.g., @banerjee2010pitfalls; @lieberman2014does;
@raffler2018weakness].

What factors, then, are driving these results? It seems unlikely that the null
results derive from flaws in the research design. First, our study is well
powered. Although the treatment is indirect---the person has to download the
app, find a school construction, and then report it---,we included
a substantial number of schools in the treatment groups. Second, balance and
manipulation tests indicate that the treatment allocation was successful, so we
can rule out problems in the randomisation procedures. Third, data from
Google Analytics confirm that citizens indeed used the app and provided
information to our dataset. This indicates that the treatment manipulation was
effective. In this sense, it is unlikely that our results derive from low user
response. Fourth, after doing a series of robustness tests, we still find no
firm evidence of treatment effect. Finally, note that the sign of the
coefficients are frequently contrary to our theoretical expectations. This
rules out a possible concern about statistical power with our small control
group approach.

We discuss some possible reasons why community monitoring did not work in our
case. One plausible explanation is that individuals were unable to
differentiate the effect of political corruption from those of spending cuts.
Due to the severe economic crisis in 2014--2016, the Brazilian federal
government introduced discretionary spending limits that affected public
investment [@rossi2016impactos]. Politicians may argue that delays in school
constructions are not derived from their misuse of government funds, but from
the austerity measures. If this is the case, citizens will not blame local
politicians for the underprovision of public goods. Consequently,
representatives can dismiss individual requests as the issue is unlikely to
escalate.

The electoral cycle might also have decreased the potential effect of the
treatment. As the experiment was fielded right after Brazil's municipal
elections, incumbents might have disregarded the requests because they did not
see the demands as politically costly in the short run. Having just taken
office, mayors might have focused their attention to the formation of
government coalitions or to budget concerns. Future research may evaluate how
electoral dynamics interact with citizen oversight, potentially by replicating
informational experiments in different stages of the political cycle.

In sum, our experiments suggest that popular participation and bottom-up
monitoring may not be effective to improve public service delivery in the case
of school constructions in Brazil. Nevertheless, the null findings are
informative to researchers and policy-makers. The most important recommendation
derived from this study is that interventions targeting elite groups, such as
lobbyists or civil servants, might render better school construction outcomes
than those focused at the community level. Another core lesson is that although
digital interventions are promising means to deliver information, perhaps they
do not have the same impact as personal, face-to-face communication. Since many
developing countries share Brazil's issues with education provision, the
shortcomings we describe here serve as warnings for future interventions.
Finally, whether this study generalises beyond school constructions to other
bottom-up programs, and to other contexts, remains to be studied.

\newpage
\setlength{\parindent}{0cm} 
\setlength{\parskip}{5pt}
